### 2021-04-04 Accessibility-Vaccination-Sites-Hamilton/Accessibility-Vaccination-Sites-Hamilton.Rmd
Initial commit with data package

---
title: "An examination of accessibility to pilot COVID-19 vaccination sites in Hamilton, Ontario"
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: 1
  - name: Christopher D. Higgins
    email: cd.higgins@utoronto.ca
    affiliation: University of Toronto Scarborough
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, McMaster University, Hamilton, ON, L8S 4K1, Canada
  - code: University of Toronto Scarborough
    address: Department of Geography & Planning, University of Toronto Scarborough, 1265 Military Trail, Toronto, ON M1C1A4
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  Abstract

journal: "Transport Findings"
date: "`r Sys.Date()`"
bibliography: References.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
always_allow_html: true

header-includes:
 \usepackage{float}
 \usepackage{xcolor}
 \floatplacement{figure}{H}
---

```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `sobiEquity`
if (!require("vaccHamilton", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/Accessibility-Pharmacies-Hamilton-Vaccines", subdir = "vaccHamilton")
  }
```

```{r load-packages, include=FALSE}
#Load the packages needed to read and work with the data
#library(circlize)
library(grid)
library(gridExtra)
library(kableExtra)
library(png)
#library(readr)
library(sf)
library(tidyverse)
library(vaccHamilton)
```

Research Questions and Hypotheses
==========================

Words.

The objective of this paper is to investigate...Using data...we pose the following questions:

- Question 1
- Question 2

This paper is a reproducible research document [see @Brunsdon2020opening]; the code and data necessary to reproduce the tables and figures are available in a public repository\footnote{\url{https://github.com/paezha/Accessibility-Pharmacies-Hamilton-Vaccines}}

Methods and Data
============

Words.

Findings
===================

Words.

References {#references .unnumbered}
==========