### 2021-04-04 README.Rmd

---
output:
  github_document:
    pandoc_args: --webtex
always_allow_html: true
bibliography: Accessibility-Vaccination-Sites-Hamilton/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# An examination of accessibility to pilot COVID-19 vaccination sites in Hamilton, Ontario

<!-- badges: start -->
<!-- badges: end -->

Antonio Páez (McMaster University)  
Christopher D. Higgins (University of Toronto Scarborough)

Paper submitted to *Transport Findings*

## Abstract

Words.

## Keywords

- COVID-19  
- Vaccination sites
- Accessibility  
- Hamilton  
- Ontario
- Reproducible research