### 2021-04-05 README.Rmd

---
output:
  github_document:
    pandoc_args: --webtex
always_allow_html: true
bibliography: Accessibility-Vaccination-Sites-Hamilton/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# An examination of the accessibility implications of a pilot COVID-19 vaccination program in Hamilton, Ontario

<!-- badges: start -->
<!-- badges: end -->

Antonio Páez (McMaster University)  
Christopher D. Higgins (University of Toronto Scarborough)

Paper submitted to *Transport Findings*

## Abstract

The province of Ontario in Canada announced the pilot for a new vaccination program, with designated pharmacies across the province now able to offer COVID-19 vaccines. The accessibility of this program raises questions about the cost of travel and the distribution of the cost among the population. In our examination of the City of Hamilton we find that selected sites do not serve well the rural and urban population of Hamilton, and that the associated cost of travel is expected to be disproportionally borne by lower income populations. Modest additions to the list of pilot sites in the city can substantially alleviate this inequity.

## Keywords

- COVID-19  
- Vaccination sites
- Accessibility  
- Hamilton  
- Ontario
- Reproducible research

```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `vaccHamilton`
if (!require("vaccHamilton", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/Accessibility-Pharmacies-Hamilton-Vaccines", 
                               subdir = "vaccHamilton")
  }
```

```{r load-packages, include=FALSE}
#Load the packages needed to read and work with the data
library(grid)
library(gridExtra)
library(kableExtra)
library(png)
library(sf)
library(tidyverse)
library(units)
library(vaccHamilton) # Data package
```

```{r invoke-data, include=FALSE}
data("data_da_2016")
data("dwelling_network_points_2016")
data("hamilton_cma")
data("modes_55to69")
data("pharmacy_locations")
data("ttm_car")
data("ttm_transit")
data("ttm_walk")
data("urban_types")
```

```{r population-age-55-plus, include=FALSE}
# Data preparation
## Calculate number of people age 55+ per DA:
data_da_2016 <- data_da_2016 %>%
  filter(`Region Name` == "Hamilton") %>%
  mutate(Population_55to69 = (`v_CA16_208: 55 to 59 years` + 
                  `v_CA16_226: 60 to 64 years` + 
                  `v_CA16_247: 65 to 69 years`),
         Population_70plus = (`v_CA16_265: 70 to 74 years` +
                  `v_CA16_283: 75 to 79 years` +
                  `v_CA16_301: 80 to 84 years` + 
                  `v_CA16_319: 85 years and over`))
```

```{r select-pharmacies, include=FALSE}
# Data preparation
## Select pharmacy locations:
pharmacy_locations <- pharmacy_locations %>%
  slice(1:23)

## Also filter selected pharmacies from travel time tables:
ttm_car <- ttm_car %>%
  dplyr::filter(OBJECTID <= 23)

ttm_transit <- ttm_transit %>%
  dplyr::filter(OBJECTID <= 23)

ttm_walk <- ttm_walk %>%
  dplyr::filter(OBJECTID <= 23)
```

```{r clip-da-geometry, include=FALSE}
# Data preparation
## This is purely cosmetic: clip `data_da_2016`:
data_da_2016 <- data_da_2016 %>%
  st_intersection(hamilton_cma)
```

```{r number-of-residential-units-per-DA, include=FALSE}
# Data preparation
## Find number of residential units in each DA. This is needed to assign the population to parcels
residential_units <- dwelling_network_points_2016 %>%
  st_drop_geometry() %>%
  group_by(DAUID) %>%
  summarize(residential_units = sum(Dwellings))
```

```{r join-population-to-parcels, include=FALSE}
# Data preparation
## Join number of people age 55+ per DA to dwellings and total number of residential units:
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID, 
                               Population_55to69,
                               Population_70plus),
            by = "DAUID") %>%
  left_join(residential_units,
            by = "DAUID")
```

```{r calculate-population-per-parcel, include=FALSE}
# Data preparation
##Calculate population per point on the network. This is the population in the DA divided by the total number of residential units. This gives the average number of people per residential unit. Then, multiply by the number of dwellings (residential units) at the point:
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  mutate(Population_55to69 = (Population_55to69/residential_units) * Dwellings,
         Population_70plus = (Population_70plus/residential_units) * Dwellings)
```

```{r join-region-to-parcels, include=FALSE}
dwelling_network_points_2016 <- dwelling_network_points_2016 %>%
  st_join(urban_types)
```

```{r calculate-population-statistics, include=FALSE}
population_statistics <- dwelling_network_points_2016 %>% 
  st_drop_geometry() %>% 
  dplyr::filter(!is.na(Type)) %>% 
  group_by(Type) %>% 
  summarize(Population_55to69 = sum(Population_55to69, 
                                    na.rm = TRUE), 
            Population_70plus = sum(Population_70plus, 
                                    na.rm = TRUE))
```

```{r join-population-data-to-travel-time-tables, include=FALSE}
# Data preparation
## Join the population data to the travel time matrices:
ttm_car <- ttm_car %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            Population_55to69,
                            Population_70plus,
                            Urban_type = Type,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            Population_55to69,
                            Population_70plus,
                            Urban_type = Type,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(ID, 
                            Population_55to69,
                            Population_70plus,
                            Urban_type = Type,
                            DAUID, 
                            TAZUID),
            by = c("UID" = "ID"))
```

```{r join-pharmacy-data-to-travel-time-tables, include=FALSE}
# Data preparation
## Join the pharmacy data to the travel time matrices:
ttm_car <- ttm_car %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))

ttm_transit <- ttm_transit %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))

ttm_walk <- ttm_walk %>%
  left_join(pharmacy_locations %>%
              st_drop_geometry() %>%
              dplyr::select(ID, 
                            Type),
            by = c("OBJECTID" = "ID"))
```

```{r drop-missing-values, include=FALSE}
# Data preparation
## Drop parcels with missing values:
ttm_car <- ttm_car %>%
  drop_na(TAZUID,
          Population_55to69)

ttm_transit <- ttm_transit %>%
  drop_na(TAZUID,
          Population_55to69)

ttm_walk <- ttm_walk %>%
  drop_na(TAZUID,
          Population_55to69)
```

```{r proportion-of-trips-by-mode, include=FALSE}
# Data preparation
## Calculate proportion of trips by mode per TAZ:
modes <- modes_55to69 %>%
  mutate(p_car = Driver/(Driver + Transit + Walk),
         p_transit = Transit/(Driver + Transit + Walk),
         p_walk = Walk/(Driver + Transit + Walk))
```

```{r complete-proportion-of-trips-by-mode, include=FALSE}
# Data preparation
## Assume that TAZ with NAs have a proportion of car trips of one:
modes <- modes %>%
  dplyr::mutate(p_car = replace_na(p_car, 1),
                p_transit = replace_na(p_transit, 0),
                p_walk = replace_na(p_walk, 0))
```

```{r join-proportion-of-trips-to-travel-time-tables, include=FALSE}
# Data preparation
## Join proportion of trips by mode to travel time tables:
ttm_car <- ttm_car %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_transit <- ttm_transit %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")

ttm_walk <- ttm_walk %>%
  left_join(modes %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        p_car,
                        p_transit,
                        p_walk),
            by = "TAZUID")
```

```{r income-quintiles, include=FALSE}
# Data preparation
## Calculate quintiles of median household income in Hamilton:
mhi_q <- data_da_2016 %>%
  dplyr::filter(`Region Name` == "Hamilton") %>%
  pull(`v_CA16_2397: Median total income of households in 2015 ($)`) %>%
  quantile(c(0, 0.2, 0.4, 0.6, 0.8, 1), 
           na.rm = TRUE)
```

```{r label-income-quintiles, include=FALSE}
# Data preparation
## Label the DAs according to income quintiles
data_da_2016 <- data_da_2016 %>%
  rename(Median_household_income = `v_CA16_2397: Median total income of households in 2015 ($)`) %>%
  mutate(income_quintile = case_when(Median_household_income < mhi_q[2] ~ "Bottom 20%",
                                     Median_household_income >= mhi_q[2] & Median_household_income < mhi_q[3] ~ "Fourth 20%",
                                     Median_household_income >= mhi_q[3] & Median_household_income < mhi_q[4] ~ "Third 20%",
                                     Median_household_income >= mhi_q[4] & Median_household_income < mhi_q[5] ~ "Second 20%",
                                     Median_household_income >= mhi_q[5] ~ "Top 20%"),
         income_quintile = factor(income_quintile,
                                  levels = c("Top 20%",
                                             "Second 20%",
                                             "Third 20%",
                                             "Fourth 20%",
                                             "Bottom 20%"),
                                  ordered = TRUE))
```

```{r minimum-travel-time-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Retrieve the minimum travel time from a point to a pharmacy by each mode for people aged 55 to 69:
min_ttm_car <- ttm_car %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_transit <- ttm_transit %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_walk <- ttm_walk %>%
  dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")
```

```{r weighted-travel-time-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Calculate weighted travel time:
min_ttm_car <- min_ttm_car %>%
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit <- min_ttm_transit %>%
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk <- min_ttm_walk %>%
  mutate(weighted_travel_time = travel_time * p_walk)
```

```{r bind-travel-time-tables-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Bind the three tables:
min_ttm <- rbind(min_ttm_car,
                 min_ttm_transit,
                 min_ttm_walk)
```

```{r aggregate-weighted-travel-times-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Aggregate weighted travel times by mode:
min_ttm <- min_ttm %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            Population = first(Population),
            Urban_type = first(Urban_type),
            weighted_travel_time = sum(weighted_travel_time))
```

```{r calculate-person-hours-travel-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Calculate person-hours of travel:
min_ttm <- min_ttm %>%
  mutate(weighted_person_hours = Population * weighted_travel_time/60)
```

```{r summarize-results-by-taz-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Summarize by TAZ and join geometry:
min_ttm_taz <- min_ttm %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            weighted_person_hours = sum(weighted_person_hours)) %>%
  left_join(modes %>%
              dplyr::select(TAZUID),
            by = "TAZUID") %>%
  st_as_sf()
```

```{r weighted-travel-time-map-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted travel time:
ttm_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1) +
  geom_sf(data = urban_types,
          aes(color = Type),
          fill = NA) +
  scale_color_manual(name = "",
                     values = c("Urban" = "firebrick2",
                                "Suburban" = "cornflowerblue",
                                "Rural" = "forestgreen"))  +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))
```

```{r person-hours-traveled-map-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted person-hours of travel:
pht_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_person_hours),
          color = NA) +
  scale_fill_distiller(name = "Person-hours of travel",
                       palette = "OrRd", 
                       direction = 1) +
  geom_sf(data = urban_types,
          aes(color = Type),
          fill = NA) +
  scale_color_manual(name = "",
                     values = c("Urban" = "firebrick2",
                                "Suburban" = "cornflowerblue",
                                "Rural" = "forestgreen"))  +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))
```

```{r join-income-information-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Join median household information to table:
min_ttm <- min_ttm %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID,
                               income_quintile),
            by = "DAUID")
```

```{r plot-distribution-travel-time-by-income-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Distribution of weighted travel time by income:
ttm_income_baseline <- ggplot(data = min_ttm, 
       aes(x = weighted_travel_time, 
           color= income_quintile)) +
  geom_density(size = 0.5, adjust = 2.5) +
  labs(x = "Travel time (in minutes; weighted by mode)") +
  scale_color_manual(values = c("Top 20%" = "blue",
                                "Second 20%" = "steelblue1",
                                "Third 20%" = "gray40", 
                                "Fourth 20%" = "tomato",
                                "Bottom 20%" = "red")) +
  theme_minimal() +
  theme(text = element_text(size = 7),
        legend.text = element_text(size = 7))
```

```{r plot-distribution-travel-time-by-region-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Distribution of weighted travel time by region:
ttm_region_baseline <- ggplot(data = min_ttm %>%
                                drop_na(Urban_type), 
       aes(x = weighted_travel_time, 
           color= Urban_type)) +
  geom_density(size = 0.5, adjust = 3) +
  labs(x = "Travel time (in minutes; weighted by mode)") +
  scale_color_manual(name = "",
                     values = c("Urban" = "firebrick2",
                                "Suburban" = "cornflowerblue",
                                "Rural" = "forestgreen"))  +
  theme_minimal() +
  theme(text = element_text(size = 7),
        legend.text = element_text(size = 7))
```

```{r total-weighted-person-hours-by-group-baseline, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Total weighted person-hours by income:
person_hours_income_baseline <- min_ttm %>%
  group_by(income_quintile) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = income_quintile)

## Total weighted person-hours by region:
person_hours_region_baseline <- min_ttm %>%
  group_by(Urban_type) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  drop_na() %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = Urban_type)

person_hours_baseline <- rbind(person_hours_income_baseline,
                               person_hours_region_baseline)
```

```{r minimum-travel-time, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Retrieve the minimum travel time from a point to a pharmacy by each mode for people aged 55 to 69:
min_ttm_car <- ttm_car %>%
  #dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_transit <- ttm_transit %>%
  #dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")

min_ttm_walk <- ttm_walk %>%
  #dplyr::filter(Type == "Pilot") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            Population = first(Population_55to69),
            Urban_type = first(Urban_type),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            .groups = "drop")
```

```{r weighted-travel-time, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Calculate weighted travel time:
min_ttm_car <- min_ttm_car %>%
  mutate(weighted_travel_time = travel_time * p_car)

min_ttm_transit <- min_ttm_transit %>%
  mutate(weighted_travel_time = travel_time * p_transit)

min_ttm_walk <- min_ttm_walk %>%
  mutate(weighted_travel_time = travel_time * p_walk)
```

```{r bind-travel-time-tables, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Bind the three tables:
min_ttm <- rbind(min_ttm_car,
                 min_ttm_transit,
                 min_ttm_walk)
```

```{r aggregate-weighted-travel-times, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Aggregate weighted travel times by mode:
min_ttm <- min_ttm %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            Population = first(Population),
            Urban_type = first(Urban_type),
            weighted_travel_time = sum(weighted_travel_time))
```

```{r calculate-person-hours-travel, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Calculate person-hours of travel:
min_ttm <- min_ttm %>%
  mutate(weighted_person_hours = Population * weighted_travel_time/60)
```

```{r summarize-results-by-taz, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Summarize by TAZ and join geometry:
min_ttm_taz <- min_ttm %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            weighted_person_hours = sum(weighted_person_hours)) %>%
  left_join(modes %>%
              dplyr::select(TAZUID),
            by = "TAZUID") %>%
  st_as_sf()
```

```{r weighted-travel-time-map, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Map weighted travel time. This is the average of the weighted travel time per TAZ:
ttm_map <- ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

```{r total-person-hours-map, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Map weighted person-hours of travel. This is the total person-hours traveled per TAZ:
pht_map <- ggplot() +
  geom_sf(data = min_ttm_taz,
          aes(fill = weighted_person_hours),
          color = NA) +
  geom_sf(data = hamilton_cma,
          fill = NA) +
  scale_fill_distiller(palette = "OrRd", 
                       direction = 1)
```

```{r join-income-information, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Join median household information to table:
min_ttm <- min_ttm %>%
  left_join(data_da_2016 %>%
              st_drop_geometry() %>%
              dplyr::transmute(DAUID = GeoUID,
                               income_quintile),
            by = "DAUID")
```

```{r plot-distribution-travel-time-by-income, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Distribution of weighted travel time by income:
ttm_income <- ggplot(data = min_ttm, 
       aes(x = weighted_travel_time, 
           color= income_quintile)) +
  geom_density(size = 0.5, adjust = 2.5) +
  labs(x = "Travel time (in minutes; weighted by mode)") +
  scale_color_manual(name = "Income quintile",
                     values = c("Top 20%" = "blue",
                                "Second 20%" = "steelblue1",
                                "Third 20%" = "gray40", 
                                "Fourth 20%" = "tomato",
                                "Bottom 20%" = "red")) +
  theme_minimal() +
  theme(text = element_text(size = 7),
        legend.text = element_text(size = 7))
```

```{r plot-distribution-travel-time-by-region, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Distribution of weighted travel time by region:
ttm_region <- ggplot(data = min_ttm %>%
                                drop_na(Urban_type), 
       aes(x = weighted_travel_time, 
           color= Urban_type)) +
  geom_density(size = 0.5, adjust = 3) +
  labs(x = "Travel time (in minutes; weighted by mode)") +
  scale_color_manual(name = "Region",
                     values = c("Urban" = "firebrick2",
                                "Suburban" = "cornflowerblue",
                                "Rural" = "forestgreen"))  +
  theme_minimal() +
  theme(text = element_text(size = 7),
        legend.text = element_text(size = 7))
```

```{r total-weighted-person-hours-by-group, include=FALSE}
# Scenario analysis (PILOT AND CANDIDATE LOCATIONS)
## Total weighted person-hours by income:
person_hours_income <- min_ttm %>%
  group_by(income_quintile) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = income_quintile)

## Total weighted person-hours by region:
person_hours_region <- min_ttm %>%
  group_by(Urban_type) %>%
  summarize(total_weighted_person_hours = sum(weighted_person_hours),
            total_population = sum(Population),
            .groups = "drop") %>%
  drop_na() %>%
  mutate(ratio = total_weighted_person_hours/total_population) %>%
  rename(group = Urban_type)

person_hours <- rbind(person_hours_income,
                      person_hours_region)
```

```{r table-comparison, echo=FALSE}
table_comparison <- cbind(person_hours_baseline %>%
        dplyr::select(group, total_population, total_weighted_person_hours, ratio), 
      person_hours %>%
        dplyr::select(total_weighted_person_hours, ratio))
```

## Research Questions and Hypotheses

Along with the provision of health care facilities to treat severe cases of COVID-19 [@Pereira2021geographic], another front in the fight against the pandemic is the rolling out of vaccination programs. The Province of Ontario, in Canada, announced on April 1st 2021 the expansion of a pilot program to offer vaccines in pharmacies in the City of Hamilton\footnote{\url{https://www.cbc.ca/news/canada/hamilton/astrazeneca-vaccine-hamilton-1.5972704}}. This program is in addition to dedicated vaccination centers for people aged 70+. Twenty pharmacies in Hamilton were added to an earlier list of 325 locations in other cities across the province, and the program was extended to people aged 55 years old and over.

Critics were swift to point out that the list of pharmacies approved for Hamilton by the province were mostly located in lower density parts of the city that are not well serviced by transit and are difficult to reach by foot\footnote{See inter alia: \url{https://twitter.com/RyanMcGreal/status/1378027149790224386?s=20} and \url{https://twitter.com/NrinderWard3/status/1378679195514060801?s=20}.}. Indeed, as seen in Figure \ref{fig:pharmacies-and-regions}, a vast majority of the pharmacies are in suburban Hamilton. The issue is somewhat less clear-cut when we consider that Hamilton's population skews suburban (see Figure \ref{fig:population-map}). Given the target demographic for the program, it is possible that suburban sites could be convenient for mature adults and young old: the population aged 55 to 69 in Hamilton is approximately `r prettyNum(round((table_comparison[7,2]/1000), 2) * 1000, big.mark = ",")` suburban, `r prettyNum(round((table_comparison[8,2]/1000), 2) * 1000, big.mark = ",")` urban, and only `r prettyNum(round((table_comparison[6,2]/1000), 2) * 1000, big.mark = ",")` rural. Nevertheless, the selection of sites by the province raises some important questions\footnote{The decision-making process to select these sites appears to have been opaque, and the typically inert Major of the city was caught flat footed by the announcement; see: \url{https://twitter.com/FredEisenberger/status/1378350123114242053?s=20}}. As @Yu2021sustained note, good geographical coverage is a key element for a successful vaccination campaign; at the same time, siting vaccinations sites in car-oriented locations may introduce inequities in access.

In this research, we investigate the accessibility implications of the sites selected for the pilot vaccination program. Concretely, we ask:

- What is the estimated cost of travel to reach the vaccination sites, assuming that every person requires a vaccine?
- What is the distribution of this cost across the population of the city?
- How does the cost and its distribution change with the addition of candidate sites in urban Hamilton?

We concentrate on the 55 to 69 years old population segment because the older 70+ group have access to other dedicated facilities besides those in the provincial pilot.

```{r pharmacies-map, echo=FALSE, fig.height=4, fig.align="center", fig.cap="\\label{fig:pharmacies-and-regions}Regions with the City of Hamilton; the location of pharmacies in pilot is shown (black triangles) and urban locations for scenario analysis (white circles)"}
#map_pharmacies <-
ggplot() +
  geom_sf(data = urban_types,
          aes(fill = Type),
          color = "lightgray") +
  geom_sf(data = pharmacy_locations,
          aes(shape = Type,
              color = Type),
          size = 2.5,
          show.legend = FALSE) +
  scale_color_manual(values = c("Candidate" = "white",
                                "Pilot" = "black")) +
  scale_fill_manual(values = c("Urban" = "firebrick2",
                               "Suburban" = "cornflowerblue",
                               "Rural" = "forestgreen")) +
  guides(fill = guide_legend(title = "")) + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))

```

```{r population-map, echo=FALSE, warning=FALSE, fig.height=8, fig.align="center", fig.cap="\\label{fig:population-map}Distribution of population age 55+ in the City of Hamilton"}
ggplot() +
  geom_sf(data = data_da_2016 %>%
            filter(`Region Name` == "Hamilton") %>%
            dplyr::select(Population_55to69, Population_70plus) %>%
            pivot_longer(cols = -geometry, names_to = "Age", values_to = "Population") %>%
            mutate(Age = factor(Age,
                                levels = c("Population_55to69",
                                           "Population_70plus"),
                                labels = c("Age: 55 to 69",
                                           "Age: 70+"),
                                ordered = TRUE)) %>%
            st_as_sf(),
          aes(fill = Population),
          color = NA) +
  geom_sf(data = urban_types,
          aes(color = Type),
          fill = NA) +
  scale_fill_distiller(name = "Population",
                       breaks = c(20, 50, 150, 400),
                       palette = "OrRd", 
                       trans = "log",
                       direction = 1) +
  scale_color_manual(name = "",
                     values = c("Urban" = "firebrick2",
                                "Suburban" = "cornflowerblue",
                                "Rural" = "forestgreen"))  +
  facet_wrap(~ Age, nrow = 2) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "bottom",
        legend.text = element_text(size = 8))
```

## Methods and Data

## Data

We use data from the following sources.

### Open Hamilton

From the open data portal of the City of Hamilton\footnote{\url{https://open.hamilton.ca/}} we obtained boundaries for the city's various regions (the definition of urban, suburban, and rural regions follows the classification of development applications).

### Statistics Canada

Population and income statistics at the level of Dissemination Areas (DAs) were retrieved using the package `cancensus` [@vonBergmann2021cancensus]. Dissemination Areas are the smallest publicly available census geography in Canada. We use data from the 2016 Population Census.

### Transportation Tomorrow Survey

Data about modal split by age by place of residence were downloaded from the Data Retrieval System of the Transportation Tomorrow Survey (TTS)\footnote{\url{http://dmg.utoronto.ca/}}. The data are geocoded at the level of Traffic Analysis Zones (TAZ).

### Other

The locations of pharmacies in the pilot were obtained from public records and geocoded. Three urban sites not in the program were also identified and geocoded for comparison purposes. In addition, we converted all recorded residential parcels in the City of Hamilton to points on the road network. Each point includes information about the number of residential units in the parcel.

## Methods

We used the population aged 55 to 59 y.o. in each Dissemination Area to calculate the average number of people per dwelling. This value was then assigned proportionally to the number of dwellings per parcel. The median total household income of the corresponding DAs was joined to the parcels. In addition, we calculated the proportion of trips by mode from the total number of trips by each of the three modes retrieved from the TTS data. These proportions were joined to the parcel data based on their corresponding TAZ. The package `r5r` [@Pereira2021r5r] was used to calculate the travel time from each parcel to all pharmacies by three modes: car, transit, and walking. For routing purposes we used a cutoff value of 180 min and a maximum walking distance of 10,000 m.

Once we obtained travel time tables with population, proportion of trips by mode, and income information, we calculated the expected travel time $ett$ from each parcel $i$ to a pharmacy $j$ as follows:
$$
ett_i = p^c_i \min(tt^c_{ij}) + p^t_i \min(tt^t_{ij}) + p^w_i \min(tt^w_{ij})
$$
\noindent where $p^k_i$ is the proportion of trips by mode $k$ in the TAZ of parcel $i$, and $tt^k_{ij}$ is the vector of travel times from parcel $i$ to the pharmacies. In other words, the expected travel time is the weighted sum of travel times to the nearest pharmacy, with the weights given by the expected modal split in the TAZ.

The expected travel time $i$ was multiplied by the population in parcel $i$ to obtain a measure of person-hours of travel ($PHT$) as follows:
$$
PHT_i = P_i\cdot ett_i
$$

Please note that this paper is a reproducible research document [see @Brunsdon2020opening] conducted using open source tools for transportation analysis [@Lovelace2021open]. The code and data necessary to reproduce the analysis are available in a public repository\footnote{\url{https://github.com/paezha/Accessibility-Pharmacies-Hamilton-Vaccines}}.

## Findings

The top panel of Figure \ref{fig:maps-baseline} shows the average expected travel time by TAZ in Hamilton. It is apparent that travel times tend to be lower in much of suburban Hamilton, and higher in the urban core and some rural parts of the city, particularly to the west. This is unsurprising, given the higher probability of travel by car and the predominantly suburban character of the vaccination sites. However, even accounting for the distribution of population, this leads to large disparities in the number of person-hours of travel across the city, with a concentration of the burden of travel in the urban core and the rural west (see bottom panel of Figure \ref{fig:maps-baseline}).

The disparities are not trivial.

As seen in Table \ref{tab:distribution-results}, under the pilot program approximately `r round((table_comparison[5,2] + table_comparison[4,2]) / (sum(table_comparison[1:5,2])) * 100, 2)`\% of people live in DAs in the bottom 40% of the median household income scale, but they account for `r round((table_comparison[5,3] + table_comparison[4,3]) / (sum(table_comparison[1:5,3])) * 100, 2)`\% of the total person-hours of travel. In contrast, `r round((table_comparison[1,2] + table_comparison[2,2]) / (sum(table_comparison[1:5,2])) * 100, 2)`\% of people aged 55 to 69 in DAs in the top 40% of the median household income scale accrue only `r round((table_comparison[1,3] + table_comparison[2,3]) / (sum(table_comparison[1:5,3])) * 100, 2)`\% of the total person-hours of travel. Where the mean travel time of residents of DAs with high median household income is `r table_comparison[1, 4] %>% set_units(h) %>% set_units(min) %>% round() %>% drop_units()` minutes, residents of lower income DAs average `r table_comparison[5, 4] %>% set_units(h) %>% set_units(min) %>% round() %>% drop_units()` minutes in travel time. In addition to longer average travel time, residents in lower income DAs also see substantially larger variations in travel times, and some may face considerably longer travel times (see top-left panel in Figure \ref{fig:results}).

There are also important disparities by region. As shown in Table \ref{tab:distribution-results}, the urban and rural populations in Hamilton are approximately `r round((table_comparison[6,2] + table_comparison[8,2])/ (sum(table_comparison[6:8,2])) * 100, 2)`\% of the population but they bear `r round((table_comparison[6,3] + table_comparison[8,3])/ (sum(table_comparison[6:8,3])) * 100, 2)`\% of the total person-hours of travel, with also much greater variability in expected travel times (Figure \ref{fig:results}, bottom-left panel).

For comparison purposes we consider a scenario with some modest additions to the list of pharmacies in the provincial pilot. We repeat the analysis, but include the three urban sites shown in white circles in Figure \ref{fig:pharmacies-and-regions}. The results of this scenario appear in the last two columns of Table \ref{tab:distribution-results} and the two right panels of Figure \ref{fig:results}. We begin by noting that all income groups benefit from the addition of these three sites with shorter mean trip durations; the most remarkable difference is the large reduction in the disparities between residents in DAs with different levels of income. The top-right panel of Figure \ref{fig:results} shows that the distribution of expected travel time is now more in line for all income groups, even if the bottom two income quintiles still have somewhat wider spreads. With respect to the distribution of expected travel time and person-hours of travel by region, unsurprisingly the addition of three urban vaccination sites in the scenario makes a difference for urban residents but not for rural residents.

The results indicate that the locations chosen by the province for the pilot vaccination program do not serve well urban or rural residents of the city, and there are some important questions regarding equity of access to the program, with a disproportionate burden in the cost of travel falling on lower income urban populations and rural populations. A scenario did not consider candidate sites in a systematic way. Nonetheless, selection of three sensible urban locations does much to alleviate disparities in the burden of transportation. On the other hand, unlike the urban context where there are numerous candidate locations that could be chosen for a vaccination program, there are not many candidate locations in rural regions of the city. Increasing access in rural Hamilton likely will involve an expansion of existing mobile vaccination pop-up clinics\footnote{\url{https://www.hamilton.ca/government-information/news-centre/news-releases/hamiltons-covid-19-vaccination-program-expansion-1}}.

```{r figure-maps-baseline, echo=FALSE, warning=FALSE, fig.height=7, fig.align="center", fig.cap="\\label{fig:maps-baseline}Average expected travel time by TAZ (in minutes) and total person-hours of travel by TAZ."}
grid.arrange(ttm_map_baseline +
               theme(legend.title = element_text(size = 8),
                     legend.text = element_text(size = 7), 
                     axis.text = element_text(size = 7)),
             pht_map_baseline +
               theme(legend.title = element_text(size = 8),
                     legend.text = element_text(size = 7), 
                     axis.text = element_text(size = 7)),
             nrow = 2)
```

```{r table-results, echo=FALSE}
table_comparison  %>%
  kable("html",
        booktabs = TRUE,
        col.names = c("Group", 
                      "Population", 
                      "Total PHT", 
                      "Hours per person", 
                      "Total PHT",
                      "Hours per person"),
        digits = 3,
        align = c("lccccc"),
        caption = "\\label{tab:distribution-results}Distribution of person-hours of travel (PHT) by median total household income and region: pilot locations only, and scenario with three urban locations added") %>% 
  add_header_above(c(" " = 2, "Pilot Program" = 2, "Scenario" = 2)) %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  pack_rows("Income Quintile", 1, 5) %>%
  pack_rows("Region", 6, 8) %>%
  footnote(general = "The population totals differ due to small differences in the classification of the regions")
```

```{r figure-results, echo=FALSE, warning=FALSE, fig.height=4, fig.align="center", fig.cap="\\label{fig:results}Distribution of expected travel time for different population groups"}
grid.arrange(ttm_income_baseline + 
               ggtitle("By income: Pilot Program") + 
               labs(x = "") + 
               xlim(c(0,40)) +
               theme(title = element_text(size = 8), 
                     axis.text = element_text(size = 7), 
                     legend.position = "none"), 
             ttm_income + 
               ggtitle("By income: Scenario") + 
               labs(x = "", y = "") + 
               xlim(c(0,40)) +
               theme(title = element_text(size = 8), 
                     axis.text = element_text(size = 7)), 
             ttm_region_baseline + 
               ggtitle("By region: Pilot Program") + 
               xlim(c(0,40)) +
               theme(title = element_text(size = 8),
                     axis.title = element_text(size = 8), 
                     axis.text = element_text(size = 7), 
                     legend.position = "none"), 
             ttm_region + 
               ggtitle("By region: Scenario") + 
               labs(y = "") +
               xlim(c(0,40)) +
               theme(title = element_text(size = 8),
                     axis.title = element_text(size = 8), 
                     axis.text = element_text(size = 7)), 
             nrow = 2)
```

## References {#references .unnumbered}